---
title: "Replicate Paper2"
author: "Yuliang Xu"
date: "May 14, 2018"
output: html_document
---
#set up
```{r setup, include=FALSE, echo=FALSE}
if(!require(frailtypack,quietly = TRUE)){
  install.packages("frailtypack")
  library(frailtypack)
}
if(!require(gtools,quietly = TRUE)){
  install.packages("gtools")
  library(gtools)
}


#setwd("M:/yuliang/project1")
#datals=read.csv('datals.csv')
datalsnew=read.csv('datalsnew.csv')
```
#data explore
```{r}
#data explore
head(datals)
dim(datals)
colnames(datals)
class(datals$t.start0)
colnames(datalsnew)
```
#shared frailty model for terminal event CRC
```{r}
#shared frailty model for terminal event CRC
CRC.SF=frailtyPenal(Surv(t_start0,t_stop0,CRC)~
                 cluster(famid)+
                 firstvisitage+
                 probandage_trans+
                 gender+
                 num_adenoma_beforecrc+
                 cnvisit_crc_10,
               data=datalsnew,
               hazard = "Splines-per",
               n.knots = 5,
               kappa = 1,
               cross.validation = TRUE)
#using @cnvisit_crc instead of @cnvisit_crc_10

```

```{r}
summary(CRC.SF)
```
#joint frailty model for terminal event CRC and recurrent polyps
```{r}
#joint frailty model for terminal event CRC and recurrent polyps

#?? how to choose the covariates?
CRC.polyps.JF = frailtyPenal(Surv(t_start0,t_stop0,polyp_event)~#polyp_event??
                               cluster(idnew)+
                               firstvisitage_trans+
                               age_prev_trans+
                               cum_adenoma_prev+#cum_adenoma_prev
                               probandage_trans+
                               gender+
                               terminal(CRC),
                             
                             formula.terminalEvent=
                               ~firstvisitage_trans+
                               num_adenoma_beforecrc+
                               probandage_trans+
                               gender,
                             
                             recurrentAG = T,
                             n.knots = 5,
                             kappa = c(6.47498, 9.74423e-19),
                             maxit = 40,
                             data=datalsnew)
```

```{r}
summary(CRC.polyps.JF)
```
#Joint nested frailty model for terminal event CRC and recurrent polyp
```{r}
#Joint nested frailty model for terminal event CRC and recurrent polyp
CRC.polyps.JNF = frailtyPenal(Surv(t_start0,t.stop0,polyp)~
                                subcluster(idnew)+
                                cluster(famid)+
                                firstvisitage_trans+
                                age_prev_trans+
                                probandage_trans+
                                gender+
                                terminal(CRC),
                              formula.terminalEvent = 
                                ~firstvisitage_trans+
                               probandage_trans+
                               gender,
                             
                            recurrentAG = T,
                            n.knots = 5,
                            kappa = c(6.47498, 9.74423e-19),
                            maxit = 40,
                            initialize = T,
                            data = datalsnew)
```

```{r}
summary(CRC.polyps.JNF)
```

#Martingale residual plots for Joint nested frailty model
```{r}
#Martingale residual plots for Joint nested frailty model
plot(aggregate(datals$t.start0,
               by=list(datals$idnew),
               FUN=max)[2][,1],
     CRC.polyps.JNF$martingale.res,
     ylab = "",
     xlab = "time",
     main = "Recurrence of polyps")

lines(lowess(aggregate(datals$t.start0,
               by=list(datals$idnew),
               FUN=max)[2][,1],
            CRC.polyps.JNF$martingale.res,
            f=1),
     lwd = 3,
     col = "red")
```
#Scatter plots for predicted individual frailty values
```{r}
#Scatter plots for predicted individual frailty values
datanew = datals[!datals$polyp==0,c('idnew')] #select polyp = T indivi
cbind(datals$idnew,datals$polyp)[1:10,]
idcount = table(factor(datanew,levels = 1:422))
idcount = as.numeric(idcount)
plot(idcount, CRC.polyps.JNF$frailty.pred,
     xlab = "Number of recurrent polyp events",
     ylab = "Individual frailty predictions for each patient",
     type = "p",
     axes = F, cex = 1.5, pch = 1)
axis(1, round(seq(0,15,length = 10), digits = 0))
axis(2, round(seq(0,15,length = 10), digits = 1))
```
#Brier score for prediction error part1
```{r}
#Brier score for prediction error part1
datasub = datalsnew
predtime = 1
window = seq(1,25,1)
fwindow = predtime + window
matsum1 = matrix(rep(0,25),nrow = 1)
#split all data into 6 folds and each with its complimentary
#the following is a huge loop
for(i in 1:6){
  if(i==1){
    datapred1 = datasub[datasub$famid %in% c(1:40),]
    datasub1 = datasub[!datasub$famid %in% c(1:40),]
  }
  if(i==2){
    datapred1 = datasub[datasub$famid %in% c(41:80),]
    datasub1 = datasub[!datasub$famid %in% c(41:80),]
  }
  if(i==3){
    datapred1 = datasub[datasub$famid %in% c(81:120),]
    datasub1 = datasub[!datasub$famid %in% c(81:120),]
  }
  if(i==4){
    datapred1 = datasub[datasub$famid %in% c(121:160),]
    datasub1 = datasub[!datasub$famid %in% c(121:160),]
  }
  if(i==5){
    datapred1 = datasub[datasub$famid %in% c(161:200),]
    datasub1 = datasub[!datasub$famid %in% c(161:200),]
  }
  if(i==6){
    datapred1 = datasub[datasub$famid %in% c(201:242),]
    datasub1 = datasub[!datasub$famid %in% c(201:242),]
  }
  #Brier score for prediction error part2
  CRC.polyps.JF5 = frailtyPenal(Surv(t_start0,t_stop0,polyp_event)~#polyp_event??
                               cluster(idnew)+
                               firstvisitage_trans+
                               age_prev_trans+
                               cum_adenoma_prev+#cum_adenoma_prev
                               probandage_trans+
                               gender+
                               terminal(CRC),
                             
                             formula.terminalEvent=
                               ~firstvisitage_trans+
                               num_adenoma_beforecrc+
                               probandage_trans+
                               gender,
                             
                             recurrentAG = T,
                             n.knots = 5,
                             kappa = c(6.47498, 9.74423e-19),
                             maxit = 40,
                             data=datasub1#different from JF
                            )
   
  #Extract the time-to-CRC
  datapred1$timetocrc = na.replace(datapred1$timetocrc,100)#replace na    with 100? why??
  timetocrc.vec = aggregate(datapred1[,c('timetocrc')],list(datapred1$idnew),mean)
  #Dynamic prediction using the training data
  #ERROR: Attempting to do predictions with a wrong model??
  #Error starts from i=2
  pred_bivariate3 = prediction(CRC.polyps.JF5,
                             data = datapred1,
                             t=predtime,
                             window = window,
                             event = "Terminal",
                             conditional = F)
  predictions_bivariate3 = pred_bivariate3$pred1
  timeto = as.vector(timetocrc.vec[,2])
  timetocrc.mat = replicate(length(window),timeto)
  fwindow.mat = t(replicate(length(timeto),fwindow))
  obs.mat = fwindow.mat > timetocrc.mat
  # Calculate Brier score
  diff.p2 = (predictions_bivariate3-obs.mat)^2
  data1 = as.data.frame(rep=i, diff.p2)
  colnames(matsum1) = c(colnames(data1))
  matsum1 = rbind.data.frame(matsum1,data1)
}
```

#Function of simulation to generate an individual for dynamic prediction
```{r}
#Function of simulation to generate an individual for dynamic prediction
#define an individual
def_person = function(gaptime,idnew=0,famid,firstvisitage,
                      adenoma_beforecrc,serrated_beforecrc,
                      others_beforecrc,gender,CRC_ind,event_indic=0){
  fistvisitage_trans = (firstvisitage-40)/10
  probandage = unique(datals$probandage[datals$famid==famid])
  probandage_trans = unique(datals$probandage_trans[datals$famid==famid])
  visitnum = ifelse(gaptime==1,10,ifelse(gaptime==3,5,ifelse(gaptime==4,4,0)))
  
  if(gaptime==1){visit = c(1,1,1,1,1,1,1,1,1,0)}
  if(gaptime==2){visit = c(1,1,1,1,1,0)}
  if(gaptime==3){visit = c(1,1,1,1,0)}
  if(gaptime==4){visit = c(1,1,1,0)}
  
  #inlude the visit before time point 10:
  if(gaptime==1){t_start0 = c(0:8,8)}
  if(gaptime==2){t_start0 = c(0,2,4,6,8,8)}
  if(gaptime==3){t_start0 = c(0,3,6,9,9)}
  if(gaptime==4){t_start0 = c(0,4,8,8)}
  
  t_stop0 = t_start0+gaptime
  # define longitudinal CRC indicator
  if(CRC_ind==1){CRC = c(rep(0,(visitnum-1)),1)}
  if(CRC_ind==0){CRC = rep(0,visitnum)}
  # define polyp type variables
  if(adenoma_beforecrc==1){
    polyp_adenoma = c(1,0,1,rep(0,(visitnum-3)))
    cum_adenoma = c(1,0,1,rep(0,visitnum-3))
    cum_adenoma_prev = c(0,1,0,1,rep(0,visitnum-4))
    num_adenoma_beforecrc = rep(2,visitnum)
  }
  if(adenoma_beforecrc==0){
    polyp_adenoma = rep(0,visitnum)
    cum_adenoma = rep(0,visitnum)
    cum_adenoma_prev = rep(0,visitnum)
    num_adenoma_beforecrc = rep(0,visitnum)
  }
  
  if(serrated_beforecrc==1){
    polyp_serrated = c(1,0,1,rep(0,(visitnum-3)))
    cum_serrated = c(1,0,1,rep(0,visitnum-3))
    cum_serrated_prev = c(0,1,0,1,rep(0,visitnum-4))
    num_serrated_beforecrc = rep(2,visitnum)
  }
  if(serrated_beforecrc==0){
    polyp_serrated = rep(0,visitnum)
    cum_serrated = rep(0,visitnum)
    cum_serrated_prev = rep(0,visitnum)
    num_serrated_beforecrc = rep(0,visitnum)
  }
  
  if(others_beforecrc==1){
    polyp_others = c(1,0,1,rep(0,(visitnum-3)))
    cum_others = c(1,0,1,rep(0,visitnum-3))
    cum_others_prev = c(0,1,0,1,rep(0,visitnum-4))
    num_others_beforecrc = rep(2,visitnum)
  }
  if(others_beforecrc==0){
    polyp_others = rep(0,visitnum)
    cum_others = rep(0,visitnum)
    cum_others_prev = rep(0,visitnum)
    num_others_beforecrc = rep(0,visitnum)
  }
  
  polyp_event = ifelse( (polyp_adenoma==1|polyp_serrated==1|polyp_others==1) ,1,0)
  num_nonade_beforecrc = num_serrated_beforecrc + num_others_beforecrc
  cum_nonade_prev = cum_serrated_prev+cum_others_prev
  
  cnvisit = c(1:(visitnum-1),visitnum-1)
  cnvisit_crc_10 = cnvisit/10
  
  #define age at visit
  if(gaptime==1){
    age_prev_tans = c((firstvisitage+(0:8)-46)/10,(firstvisitage+8-46)/10)
  }
  if(gaptime==2){
    age_prev_tans = (firstvisitage+c(0,2,4,6,8,8)-46)/10
  }
  if(gaptime==3){
    age_prev_tans = (firstvisitage+c(0,3,6,9,9)-46)/10
  }
  if(gaptime==4){
    age_prev_tans = (firstvisitage+c(0,4,8,8)-46)/10
  }
  
}

# create one person

# for ( i in 1:visitnum){
#   test.data[i,1] = 
# }

# combine with original data as follows:
datnew = rbind(test.data, datals_part)

```

